---
title: "miPEP_H_16S"
author: "Harriet"
date: "03/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note: "root" samples (roots, rhizoplane and close rhizosphere) and "soil" samples (distant rhizosphere and unplanted soils) shouldn't be analyzed together. The DNA of these 2 types of samples was not extracted in the same manner: Nucleospin Plant Kit and Griffiths protocole, respectively. Moreover, soil samples 16S and ITS were quantified by qPCR with the addition of BSA (0,67mg/mL), which was not added for the root samples. There should be 96 samples = 8 plants (RACINE) "ROOTS and close rhizosphere" * 4 conditions (miPEPa,b,c, control) + 8 soils (SOL) "distant rhizosphere" + 8 sols nus (SOL.NU) "unplanted soils". 

Loading all the necessary packages and calling my files: I'm using the otu_table_filtered.biom file which is the result of pre-processing (quality control, filtering...) done my the automatic pipeline.
ASVs found in less than 3 samples are discarded.

```{r}
library(tidyverse)
library(phyloseq)
library(data.table)
library(vegan)
library(ggplot2)
library(readxl)
library(here)
library(microbiome)

biom_file<-import_biom(here("data","16S", "otu_table_filtered.biom"), here("data","16S","tree.fasttree"), here("data","16S","otus.fasta"), parseFunction=parse_taxonomy_greengenes)
biom_file
```

A data frame is created from the mapping_file, which contains all information concerning each sample. Each row corresponds to a sample and is named in consequence. Merge the biom file and the metadata into one phyloseq object.

```{r}
sample<- data.frame(fread(here("data","16S", "mapping_file.tsv"), sep="\t"), check.names=F)
head(sample)
sample=sample_data(sample)

rownames(sample)<-sample$`#SampleID`
merge<-merge_phyloseq(biom_file,sample)
merge
```

Extract the OTU and taxa table

```{r}
tax_table = tax_table(merge)
head(tax_table)
#remplacer les noms de colonnes Rank1, rank2.. par les noms de taxons
colnames(tax_table) = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species", "rank8", "rank9", "rank10", "rank11","rank12")
head(tax_table)
#write.table(tax_table, "tax_table.txt")

OTU_table = otu_table(merge)
#write.table(OTU_table, "OTU_table.txt")
```

Count the number of sequences in each samples. 
raremax gives the sample with the lowest sequence number (usefull if rarefying)

```{r}
Seq<-sample_sums(merge)
Seq
#write.table(Seq, "sequences.txt")

min(Seq)
```

Keep ASVs that are at least present once in three samples.

```{r}
phylo_filtered = filter_taxa(merge, function(x) sum(x >= 1)>= (3), TRUE)
OTU_table_phylo_filtered = otu_table(phylo_filtered)
#write.table(OTU_table_phylo_filtered, "OTU_table_phylo_filtered.txt")
tax_table_phylo_filtered = tax_table(phylo_filtered)
#write.table(tax_table_phylo_filtered, "tax_table_phylo_filtered.txt")
```

After this filtration, you can lose a lot of ASVs (=taxa), but don't lose so much sequences. You can check that by counting them again : here we have lost 35792 sequences and 3353 bacterial taxa.

```{r}
Seq_filtered<-sample_sums(phylo_filtered)
Seq_filtered
#write.table(Seq_filtered, "sequences_filtered.txt")
```

Have a look at your sequencing depth: do all samples reach a plateau? At which sample size?
```{r}
#remotes::install_github("mahendra-mariadassou/phyloseq-extended", ref = "dev")
library(phyloseq.extended)

p<-ggrare(phylo_filtered, step=100, color = "Compartment", label = "X.SampleID", se = FALSE)
p<-p + labs(title="Rarefaction curve", subtitle =  "miPEP_H 16S samples")
p<-p+scale_color_manual(values = c("dark blue", "dark orange", "dark red"))
p<-p+theme_minimal()
p <- p + facet_wrap(~Type) 
p
```

16S/ITS sequencing doesn't allow for "real" quantitative data. To overcome this, 16S and ITS genes were quantified by qPCR, along with a standard curve, to evaluate the "absolute" number of copies of 16S and ITS in each sample. The mean Ct was converted to a number of copies, by using the equation of the standard curve. For ITS, a ratio of fungal ITS/non-fungal ITS was applied to each sample, corresponding to the sequences counted during sequencing (=fungal ASVs/total ITS ASVs). The number of copies was used to replace the total sum of 16S ASVs, per sample, and each ASV was re-caculated as a number of copies. The resulting "absolute" abundance ASV table needs to be imported into R and processed into a new phyloseq object. 
1) Export phylo_filtered OTU table 
2) Open OTU table with Excel
    - align samples with their correct column
    - add "calculs" sheet and add "Samples" line and a "ASV ID" column
    - line 2 = number of ITS copies/sample (final number of copies, after multplied by ITS ratio=nb fungal ITS/total nb of ITS)
    - CALCULATION: to repeat for each sample: ASV = nb seq ASV * nb copies ITS / sum sequences in that sample 
3) Add a spread sheed to the Excel file "quant abs"
    -copy/paste the "calculs" sheet with just the data "1,2,3"
    -delete the line 2 (with the number of copies)
    -save as an Excel file (.xlsx)
4) Put all the cells into format "Number" with no decimals and export the active sheet as .txt (Tabulation)
5) Go to Galaxy.genouest.org and upload your .txt file, then search for "convert between BIOM table formats" input =.txt and output = BIOM1 file "OTU TABLE". change nothing else 
6) Dowload the biom file and import to R  

```{r}
otu_biom <- import_biom(here("data", "16S", "otu_table_absolute_16S.biom1"))

tax <- tax_table(tax_table_phylo_filtered) #taxa hasn't changed

samples <- sample_data(merge) #use your metadata from before

phylo_abs_16S <- merge_phyloseq(tax,otu_biom,samples)

#Before continuing, we must discard the sample cR7 from the root samples, which has much less copies.
Sample_toRemove <- c("X.SampleID"="cR7")
subset_samples(phylo_abs_16S, X.SampleID %in% Sample_toRemove)
subset_samples(phylo_abs_16S, !(X.SampleID %in% Sample_toRemove))
phylo_abs_16S <- subset_samples(phylo_abs_16S, !(X.SampleID %in% Sample_toRemove))
phylo_abs_16S #should have only 95 samples now (-cR7)

OTU_table = otu_table(phylo_abs_16S)
samples_final <- sample_data(phylo_abs_16S)

phylo_abs_16S <- merge_phyloseq(tax_table,OTU_table,samples_final)

```

```{r}
correct.order <- c("miPEP-A", "miPEP-B", "miPEP-C", "Control-mi-PEP")
sample_data(phylo_abs_16S)$Type <- factor(sample_data(phylo_abs_16S)$Type, levels = correct.order)
levels(get_variable(phylo_abs_16S, "Type")) #mettre dans le bon ordre

phylo_roots <- subset_samples(phylo_abs_16S, Compartment=="RACINE")
phylo_soil <- subset_samples(phylo_abs_16S, Compartment=="SOL")
phylo_unplanted <- subset_samples(phylo_abs_16S, Compartment=="SOL.NU")
```

Venn-Diagrams to show shared ASVs between treatments: limited to 5

```{r}
# 1) Use a phyloseq object with 1 variable left, you could have to start with phylo_roots, for ex
phylo_roots

# 2) The next steps are performed on the means of each treatment, so merge treated-samples 
merge_phyloseq_subset<-merge_samples(phylo_roots,"Type")

# The Types will become numbers in the same order as in the phyloseq:
# 1=ago; 2=dcl; 3=hen; 4=RTL1; 5=RTL1myc and 6=WT

# 3) Create a table with a dataframe such as OTUs=columns and Type=rows
table<-t(merge_phyloseq_subset@otu_table@.Data)

# 4) Create a vector per Type to have the right format for Venn Diagram
cat_1 <- rownames(table)[which(table[,1]!=0)]
cat_2 <- rownames(table)[which(table[,2]!=0)]
cat_3 <- rownames(table)[which(table[,3]!=0)]
cat_4 <- rownames(table)[which(table[,4]!=0)]


# 5) These vectors need to be formatted into a list object
input <- list("miPEP-A"=cat_1,  "miPEP-B"=cat_2, "miPEP-C"=cat_3, "Control-mi-PEP"=cat_4)

library(VennDiagram)
library(gplots)
library(reshape2)
display_venn <- function(x, ...){
  library(VennDiagram)
  grid.newpage()
  venn_object <- venn.diagram(x, filename = NULL, ...)
  grid.draw(venn_object)
}
library(RColorBrewer)

myCol <- brewer.pal(4, "Set2") # the first number = nb of conditions
display_venn(input, lwd = 2, lty ='blank',  fill=myCol, # Numbers
        cex = .9,
        fontface = "italic",
        # Set names
        cat.cex = 1,
        cat.fontface = "bold",
        cat.default.pos = "outer")
```

```{r}
# 1) Use a phyloseq object with 1 variable left, you could have to start with phylo_roots, for ex
phylo_soil

# 2) The next steps are performed on the means of each treatment, so merge treated-samples 
merge_phyloseq_subset<-merge_samples(phylo_soil,"Type")

# The Types will become numbers in the same order as in the phyloseq:
# 1=ago; 2=dcl; 3=hen; 4=RTL1; 5=RTL1myc and 6=WT

# 3) Create a table with a dataframe such as OTUs=columns and Type=rows
table<-t(merge_phyloseq_subset@otu_table@.Data)

# 4) Create a vector per Type to have the right format for Venn Diagram
cat_1 <- rownames(table)[which(table[,1]!=0)]
cat_2 <- rownames(table)[which(table[,2]!=0)]
cat_3 <- rownames(table)[which(table[,3]!=0)]
cat_4 <- rownames(table)[which(table[,4]!=0)]


# 5) These vectors need to be formatted into a list object
input <- list("miPEP-A"=cat_1,  "miPEP-B"=cat_2, "miPEP-C"=cat_3, "Control-mi-PEP"=cat_4)

library(VennDiagram)
library(gplots)
library(reshape2)
display_venn <- function(x, ...){
  library(VennDiagram)
  grid.newpage()
  venn_object <- venn.diagram(x, filename = NULL, ...)
  grid.draw(venn_object)
}
library(RColorBrewer)

myCol <- brewer.pal(4, "Set2") # the first number = nb of conditions
display_venn(input, lwd = 2, lty ='blank',  fill=myCol, # Numbers
        cex = .9,
        fontface = "italic",
        # Set names
        cat.cex = 1,
        cat.fontface = "bold",
        cat.default.pos = "outer")
```

```{r}
# 1) Use a phyloseq object with 1 variable left, you could have to start with phylo_roots, for ex
phylo_unplanted

# 2) The next steps are performed on the means of each treatment, so merge treated-samples 
merge_phyloseq_subset<-merge_samples(phylo_unplanted,"Type")

# The Types will become numbers in the same order as in the phyloseq:
# 1=ago; 2=dcl; 3=hen; 4=RTL1; 5=RTL1myc and 6=WT

# 3) Create a table with a dataframe such as OTUs=columns and Type=rows
table<-t(merge_phyloseq_subset@otu_table@.Data)

# 4) Create a vector per Type to have the right format for Venn Diagram
cat_1 <- rownames(table)[which(table[,1]!=0)]
cat_2 <- rownames(table)[which(table[,2]!=0)]
cat_3 <- rownames(table)[which(table[,3]!=0)]
cat_4 <- rownames(table)[which(table[,4]!=0)]

# 5) These vectors need to be formatted into a list object
input <- list("miPEP-A"=cat_1,  "miPEP-B"=cat_2, "miPEP-C"=cat_3, "Control-mi-PEP"=cat_4)

library(VennDiagram)
library(gplots)
library(reshape2)
display_venn <- function(x, ...){
  library(VennDiagram)
  grid.newpage()
  venn_object <- venn.diagram(x, filename = NULL, ...)
  grid.draw(venn_object)
}
library(RColorBrewer)

myCol <- brewer.pal(4, "Set2") # the first number = nb of conditions
display_venn(input, lwd = 2, lty ='blank',  fill=myCol, # Numbers
        cex = .9,
        fontface = "italic",
        # Set names
        cat.cex = 1,
        cat.fontface = "bold",
        cat.default.pos = "outer")
```

Nice Theme
```{r}
  niceTheme <- theme(
    axis.text.x = element_blank(),
    axis.title.x=element_blank(),
        axis.ticks.x=element_blank(),
    axis.text.y = element_text(size = 10),
    legend.text = element_text(size = 9),
    strip.text = element_text(size = 10), strip.background = element_rect(colour = "black", fill = "white"),
  legend.position="right",panel.background = element_rect(fill = "white",colour="grey"),panel.grid.major = element_line(colour = "white"))
```


Abundance plot in all samples
```{r}
library(microbiome)
library(dplyr)
pseq <- phylo_abs_16S %>% aggregate_taxa(level = "Phylum")
#ps1.com.fam <- microbiome::aggregate_top_taxa(pseq, "Phylum", top = 10)

ps1.com.fam <- names(sort(taxa_sums(pseq), decreasing = TRUE))[1:10]
ps1.com.fam <- prune_taxa(ps1.com.fam, pseq)

plot_bar(ps1.com.fam, "Phylum", fill="Phylum", facet_grid=Compartment~Type) + niceTheme
```

Abudance plots (top phyla & genera) in roots and close rhizosphere samples
```{r}
pseq_roots <- phylo_roots %>% aggregate_taxa(level = "Phylum")
#ps1.com.fam_roots <- microbiome::aggregate_top_taxa(pseq_roots, "Phylum", top = 10)

ps1.com.fam_roots <- names(sort(taxa_sums(pseq_roots), decreasing = TRUE))[1:10]
ps1.com.fam_roots <- prune_taxa(ps1.com.fam_roots, pseq_roots)

taxa_abundance_table_phylum_roots <- psmelt(ps1.com.fam_roots)

StackedBarPlot_phylum <- taxa_abundance_table_phylum_roots %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Absolute Abundance",
       title = "Phylum Absolute Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  ) + niceTheme + scale_fill_manual(values = c("darkblue", "darkgoldenrod1","darkgrey", "darkseagreen", "darkorchid", "darkolivegreen1", "lightskyblue", "darkgreen", "deeppink", "khaki2", "firebrick", "brown1", "darkorange1", "cyan1", "royalblue4", "darksalmon", "darkblue",
"royalblue4", "dodgerblue3", "steelblue1", "lightskyblue", "darkseagreen", "darkgoldenrod1", "darkseagreen", "darkorchid", "darkolivegreen1", "brown1", "darkorange1", "cyan1", "darkgrey"))
StackedBarPlot_phylum

#Create a boxplot to show differences in composition of each phylum per Type

phyloseq::psmelt(ps1.com.fam_roots) %>%
ggplot(data = ., aes(x = Type, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = OTU), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ OTU, scales = "free") + guides(col = FALSE) +theme_classic() +theme(axis.text.x=element_text(size=8, angle=45,hjust=1)) + ggtitle(label = "Abundance of 10 bacterial phyla in roots and close rhizosphere")

# Transform Taxa counts to relative abundance for nice abundance plots
ps1_roots_phylum_relabun <- transform_sample_counts(ps1.com.fam_roots, function(OTU) OTU/sum(OTU) * 100)

taxa_abundance_table_phylum_roots_rel <- psmelt(ps1_roots_phylum_relabun)

StackedBarPlot_phylum_rel <- taxa_abundance_table_phylum_roots_rel %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Relative Abundance",
       title = "Phylum Relative Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  ) + niceTheme+
scale_fill_manual(values = c("darkblue", "darkgoldenrod1","darkgrey", "darkseagreen", "darkorchid", "darkolivegreen1", "lightskyblue", "darkgreen", "deeppink", "khaki2", "firebrick", "brown1", "darkorange1", "cyan1", "royalblue4", "darksalmon", "darkblue",
"royalblue4", "dodgerblue3", "steelblue1", "lightskyblue", "darkseagreen", "darkgoldenrod1", "darkseagreen", "darkorchid", "darkolivegreen1", "brown1", "darkorange1", "cyan1", "darkgrey"))
StackedBarPlot_phylum_rel

############# GENUS ##############

pseq_roots_genus <- phylo_roots %>% aggregate_taxa(level = "Genus")
#ps1._roots_genus_top10 <- microbiome::aggregate_top_taxa(pseq_roots_genus, "Genus", top = 10)

ps1._roots_genus_top10 <- names(sort(taxa_sums(pseq_roots_genus), decreasing = TRUE))[1:10]
ps1._roots_genus_top10 <- prune_taxa(ps1._roots_genus_top10, pseq_roots_genus)

taxa_abundance_table_genus_roots <- psmelt(ps1._roots_genus_top10)

StackedBarPlot_genus_roots <- taxa_abundance_table_genus_roots %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Genus)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Absolute Abundance",
       title = "Genus Absolute Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  ) + niceTheme+
scale_fill_manual(values = c("darkblue", "darkgoldenrod1","darkgrey", "darkseagreen", "darkorchid", "darkolivegreen1", "lightskyblue", "darkgreen", "deeppink", "khaki2", "firebrick", "brown1", "darkorange1", "cyan1", "royalblue4", "darksalmon", "darkblue",
"royalblue4", "dodgerblue3", "steelblue1", "lightskyblue", "darkseagreen", "darkgoldenrod1", "darkseagreen", "darkorchid", "darkolivegreen1", "brown1", "darkorange1", "cyan1", "darkgrey"))

StackedBarPlot_genus_roots


# Transform Taxa counts to relative abundance for nice abundance plots
genus_roots_relative <- transform_sample_counts(ps1._roots_genus_top10, function(OTU) OTU/sum(OTU) * 100)
taxa_abundance_table_genus_roots_rel <- psmelt(genus_roots_relative)

StackedBarPlot_genus_rel <- taxa_abundance_table_genus_roots_rel %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Genus)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Relative Abundance",
       title = "Genus Relative Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  ) + niceTheme + scale_fill_manual(values = c("darkblue", "darkgoldenrod1","darkgrey", "darkseagreen", "darkorchid", "darkolivegreen1", "lightskyblue", "darkgreen", "deeppink", "khaki2", "firebrick", "brown1", "darkorange1", "cyan1", "royalblue4", "darksalmon", "darkblue",
"royalblue4", "dodgerblue3", "steelblue1", "lightskyblue", "darkseagreen", "darkgoldenrod1", "darkseagreen", "darkorchid", "darkolivegreen1", "brown1", "darkorange1", "cyan1", "darkgrey"))

StackedBarPlot_genus_rel

#Boxplot

phyloseq::psmelt(ps1._roots_genus_top10) %>%
ggplot(data = ., aes(x = Type, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = OTU), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ OTU, scales = "free")  + guides(col = FALSE) +theme_classic() +theme(axis.text.x=element_text(size=8, angle=45,hjust=1)) + ggtitle(label = "Abundance of bacterial genera in roots and close rhizosphere")

```

Abudance plots (top phyla & genera) in distant rhizosphere samples
```{r}
pseq_soil <- phylo_soil %>% aggregate_taxa(level = "Phylum")
#ps1.com.fam_soil <- microbiome::aggregate_top_taxa(pseq_soil, "Phylum", top = 10)

ps1.com.fam_soil <- names(sort(taxa_sums(pseq_soil), decreasing = TRUE))[1:10]
ps1.com.fam_soil <- prune_taxa(ps1.com.fam_soil, pseq_soil)

taxa_abundance_table_phylum_soil <- psmelt(ps1.com.fam_soil)

StackedBarPlot_phylum <- taxa_abundance_table_phylum_soil %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Absolute Abundance",
       title = "Phylum Absolute Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )
StackedBarPlot_phylum

#Create a boxplot to show differences in composition of each phylum per Type

phyloseq::psmelt(ps1.com.fam_soil) %>%
ggplot(data = ., aes(x = Type, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = OTU), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ OTU, scales = "free") + guides(col = FALSE) +theme_classic() +theme(axis.text.x=element_text(size=8, angle=45,hjust=1)) + ggtitle(label = "Abundance of top bacterial phyla in distant rhizosphere")

# Transform Taxa counts to relative abundance for nice abundance plots
ps1_soil_phylum_relabun <- transform_sample_counts(ps1.com.fam_soil, function(OTU) OTU/sum(OTU) * 100)
taxa_abundance_table_phylum_soil_rel <- psmelt(ps1_soil_phylum_relabun)

StackedBarPlot_phylum_rel <- taxa_abundance_table_phylum_soil_rel %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Relative Abundance",
       title = "Phylum Relative Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )
StackedBarPlot_phylum_rel

################# GENUS #######################

pseq_soil_genus <- phylo_soil %>% aggregate_taxa(level = "Genus")
#ps1._soil_genus_top10 <- microbiome::aggregate_top_taxa(pseq_soil_genus, "Genus", top = 10)

ps1._soil_genus_top10 <- names(sort(taxa_sums(pseq_soil_genus), decreasing = TRUE))[1:10]
ps1._soil_genus_top10 <- prune_taxa(ps1._soil_genus_top10, pseq_soil_genus)

taxa_abundance_table_genus_soil <- psmelt(ps1._soil_genus_top10)

StackedBarPlot_genus_soil <- taxa_abundance_table_genus_soil %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Genus)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Relative Abundance",
       title = "Genus Relative Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

StackedBarPlot_genus_soil

#Boxplot

phyloseq::psmelt(ps1._soil_genus_top10) %>%
ggplot(data = ., aes(x = Type, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = OTU), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ OTU, scales = "free") + guides(col = FALSE) +theme_classic() +theme(axis.text.x=element_text(size=8, angle=45,hjust=1)) + ggtitle(label = "Abundance of bacterial genera in distant rhizosphere")
```

Abudance plots (top phyla & genera) in unplanted samples
```{r}
pseq_unplanted <- phylo_unplanted %>% aggregate_taxa(level = "Phylum")
#ps1.com.fam_unplanted <- microbiome::aggregate_top_taxa(pseq_unplanted, "Phylum", top = 10)

ps1.com.fam_unplanted <- names(sort(taxa_sums(pseq_unplanted), decreasing = TRUE))[1:10]
ps1.com.fam_unplanted <- prune_taxa(ps1.com.fam_unplanted, pseq_unplanted)

taxa_abundance_table_phylum_unplanted <- psmelt(ps1.com.fam_unplanted)

StackedBarPlot_phylum <- taxa_abundance_table_phylum_unplanted %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Absolute Abundance",
       title = "Phylum Absolute Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )
StackedBarPlot_phylum

#Create a boxplot to show differences in composition of each phylum per Type

phyloseq::psmelt(ps1.com.fam_unplanted) %>%
ggplot(data = ., aes(x = Type, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = OTU), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ OTU, scales = "free") + guides(col = FALSE) +theme_classic() +theme(axis.text.x=element_text(size=8, angle=45,hjust=1)) + ggtitle(label = "Abundance of top bacterial phyla in unplanted soils")

# Transform Taxa counts to relative abundance for nice abundance plots
ps1_unplanted_phylum_relabun <- transform_sample_counts(ps1.com.fam_unplanted, function(OTU) OTU/sum(OTU) * 100)
taxa_abundance_table_phylum_unplanted_rel <- psmelt(ps1_unplanted_phylum_relabun)

StackedBarPlot_phylum_rel <- taxa_abundance_table_phylum_unplanted_rel %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Phylum)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Relative Abundance",
       title = "Phylum Relative Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )
StackedBarPlot_phylum_rel

#################### GENUS #####################

pseq_unplanted_genus <- phylo_unplanted %>% aggregate_taxa(level = "Genus")
#ps1._unplanted_genus_top10 <- microbiome::aggregate_top_taxa(pseq_unplanted_genus, "Genus", top = 10)

ps1._unplanted_genus_top10 <- names(sort(taxa_sums(pseq_unplanted_genus), decreasing = TRUE))[1:10]
ps1._unplanted_genus_top10 <- prune_taxa(ps1._unplanted_genus_top10, pseq_unplanted_genus)

taxa_abundance_table_genus_unplanted <- psmelt(ps1._unplanted_genus_top10)

StackedBarPlot_genus_unplanted <- taxa_abundance_table_genus_unplanted %>% 
  ggplot(aes(x =Sample, y = Abundance, fill = Genus)) +
  geom_bar(stat = "identity") +
  labs(x = "",
       y = "Relative Abundance",
       title = "Genus Relative Abundance") +
  facet_grid(~ Type, scales = "free") +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 10),
    strip.text = element_text(size = 12)
  )

StackedBarPlot_genus_unplanted

#Boxplot

phyloseq::psmelt(ps1._unplanted_genus_top10) %>%
ggplot(data = ., aes(x = Type, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = OTU), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ OTU, scales = "free") + guides(col = FALSE) +theme_classic() +theme(axis.text.x=element_text(size=8, angle=45,hjust=1)) + ggtitle(label = "Abundance of bacterial genera in unplanted soils")
```


What is the diversity within samples like? --> calculate alpha-diversity

The following is much inspired by https://www.nicholas-ollberding.com/post/introduction-to-the-statistical-analysis-of-microbiome-data-in-r/

The measurement of alpha-diversity is never 100% accurate, as (1) richness is always underestimated due to inexhaustive sampling and (2) rare taxa cannot be accounted for, as with NGS, singletons are eliminated from the beginning, as they could be just sequencing errors.
In order to compare the index between samples (they will have similar error rates) and so that sequencing depth doesn't interfere, it is recommended to rarefy the data.
Rarefaction is the subsampling of your reads without replacement to a same sequencing depth (=number of reads).

Before rarefying, plotting Total Reads VS Observed Richness can be interesting. It seems logical that the more reads in a sample, the more potential diversity can be observed.  

In the following paragraphs, the term "reads" is used, but it is really referring to "copies", as reads have been converted to copies, in the otu_biom.

```{r}
ggplot(data = data.frame("total_reads" =  phyloseq::sample_sums(phylo_abs_16S),
"observed" = phyloseq::estimate_richness(phylo_abs_16S, measures = "Observed")[, 1]),
aes(x = total_reads, y = observed)) + geom_point() + geom_smooth(method="lm", se = FALSE)+ labs(x = "\nTotal Reads", y = "Observed Richness\n")

#Draw the rarefaction curve for phylo_abs_16S

#remotes::install_github("mahendra-mariadassou/phyloseq-extended", ref = "dev")
library(phyloseq.extended)

p<-ggrare(phylo_abs_16S, step=100, color = "Compartment", label = "X.SampleID", se = FALSE)
p<-p + labs(title="Rarefaction curve", subtitle =  "miPEP 16S samples")
p<-p+scale_color_manual(values = c("dark blue", "dark orange", "dark red"))
p<-p+theme_minimal()
p <- p + facet_wrap(~Compartment) 
p
```

Rarefaction: soil
```{r}
reads_soil <- sample_sums(phylo_soil)
min_reads_soil <- min(reads_soil)
min_reads_soil
#[1] 326 396 corresponds to aS2, the next lowest is at 588 824 aS4: this will be the rarefaction limit, to not loose any more samples. The higest sample are up to nearly 4M.

testit <- function(x)
{
    p1 <- proc.time()
    Sys.sleep(x)
    proc.time() - p1 # The cpu usage should be negligible
}
testit(10)

phylo_rar_soil <- phyloseq::rarefy_even_depth(phylo_soil, sample.size = 588824 , rngseed = 711, replace = FALSE)
```

Rarefaction roots
```{r}
reads_roots <- sample_sums(phylo_roots)
min_reads_roots <- min(reads_roots)
min_reads_roots
#[1] 67 419 corresponds to cR7, the next lowest is 4 614 100 (aR1) so it will be the rarefaction limit. The highest being near 14M.

testit(10)

phylo_rar_roots <- phyloseq::rarefy_even_depth(phylo_roots, sample.size = 4614100, rngseed = 711, replace = FALSE)

```


Rarefaction unplanted
```{r}
reads_unplanted <- sample_sums(phylo_unplanted)
min_reads_unplanted <- min(reads_unplanted)
min_reads_unplanted
# 138 478 = ctrlSN2. Next lowest :323 104, which will be the rar limit.

testit(10)

phylo_rar_unplanted <- phyloseq::rarefy_even_depth(phylo_unplanted, sample.size = 323104, rngseed = 711, replace = FALSE)

testit(5)
```

Rarefying to a certain sample.size, is often reducing sequencing depth to the smallest sample. But if some samples are really small (low number of reads), it can be better to discard those samples and rarefy at a higher size. "rngseed" allows to set a seed, meaning that this code is reproducible (normally the random rarefaction will change everytime the code is run!). "replace=FALSE" means that when a read is sampled it is not replaced in the pool of reads (https://www.thoughtco.com/sampling-with-or-without-replacement-3126563).

sometimes the rarefaction process takes up too much memory: restart R and only run the essential scripts prior to rarefaction and then do it

Alpha-diversity in root samples ...
```{r}
p_Div_roots=plot_richness(phylo_rar_roots, color="Type", x="Type", measures=c("Observed", "Shannon", "Simpson"))

p_Div_roots_bp=p_Div_roots+geom_boxplot(aes(color=Type), alpha=0.2)+ theme_classic()+xlab("Alpha-diversity in roots and close rhizosphere")+ylab("Observed ASVs richness")+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))

p_Div_roots_bp

# en noir et blanc avec des formes
p_Div_roots=plot_richness(phylo_rar_roots, shape="Type", x="Type", measures=c("Observed", "Shannon", "Simpson"))
p_Div_roots_bp_nb <- p_Div_roots+geom_boxplot(aes(shape=Type), alpha=0.2) + scale_shape_manual(values=c(0,1,2,3,19,15,8))+ theme_classic()+xlab("Alpha-diversity in roots and close rhizosphere samples")+ylab("Observed ASVs richness")+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))
p_Div_roots_bp_nb

#Statistical analyses: be careful, if updating code: this part has to be manually (export/import)

alpha.diversity_roots=estimate_richness(phylo_rar_roots, measures=c("Observed","Shannon", "Simpson"))

#write.table(alpha.diversity_roots, ("indice_diversite_roots.txt"))

#Add a "Type" column and re-import as Text(readr)

library(readr)
indice_diversite_roots <-read_delim("C:/Users/hmiddleton/Dropbox/Projet_Mutants_miPEPs_miRNA/Harriet/Metabarcoding_mutants_and_miPEP/Analyses_métabarcoding/miPEP_H/16S/indice_diversite_roots_1.txt","\t", escape_double = FALSE, trim_ws = TRUE)

indice_diversite_roots$Type <- as.factor(indice_diversite_roots$Type)
str(object = indice_diversite_roots$Type)
indice_diversite_roots$Type <- relevel(indice_diversite_roots$Type, ref="Control-mi-PEP")

fit1 <-lm(Observed~Type, data=indice_diversite_roots)
hist(fit1$residuals) #suit une loi normale?
shapiro.test(fit1$residuals)# si p-value > 0,05 alors la distribution n'est pas significativement différente d'une distribution normale
summary(fit1)

fit2 <-lm(Shannon~Type, data=indice_diversite_roots)
hist(fit2$residuals) #suit une loi normale?
shapiro.test(fit2$residuals)
summary(fit2)

fit3 <-lm(Simpson~Type, data=indice_diversite_roots)
hist(fit3$residuals) #suit une loi normale? if not normal try a GLM
shapiro.test(fit3$residuals)
glm_fit3 <- glm(Simpson~Type, data=indice_diversite_roots) #link="inverse" by default
summary(glm_fit3)
```

Alpha-diversity in soil samples ...
```{r}
p_Div_soil=plot_richness(phylo_rar_soil, color="Type", x="Type", measures=c("Observed", "Shannon", "Simpson"))

p_Div_soil_bp=p_Div_soil+geom_boxplot(aes(color=Type), alpha=0.2)+ theme_classic()+xlab("Alpha-diversity in distant rhizosphere")+ylab("Observed ASVs richness")+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))

p_Div_soil_bp

# en noir et blanc avec des formes
p_Div_soil=plot_richness(phylo_rar_soil, shape="Type", x="Type", measures=c("Observed", "Shannon", "Simpson"))
p_Div_soil_bp_nb <- p_Div_soil+geom_boxplot(aes(shape=Type), alpha=0.2) + scale_shape_manual(values=c(0,1,2,3,19,15,8))+ theme_classic()+xlab("Alpha-diversity in distant rhizosphere samples")+ylab("Observed ASVs richness")+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))
p_Div_soil_bp_nb

#Statistical analyses: be careful, if updating code: this part has to be manually (export/import)

alpha.diversity_soil=estimate_richness(phylo_rar_soil, measures=c("Observed","Shannon", "Simpson"))

#write.table(alpha.diversity_soil, ("indice_diversite_soil.txt"))

#Add a "Type" column and re-import as Text(readr)

library(readr)
indice_diversite_soil <-read_delim("C:/Users/hmiddleton/Dropbox/Projet_Mutants_miPEPs_miRNA/Harriet/Metabarcoding_mutants_and_miPEP/Analyses_métabarcoding/miPEP_H/16S/indice_diversite_soil_1.txt","\t", escape_double = FALSE, trim_ws = TRUE)

indice_diversite_soil$Type <- as.factor(indice_diversite_soil$Type)
str(object = indice_diversite_soil$Type)
indice_diversite_soil$Type <- relevel(indice_diversite_soil$Type, ref="Control-mi-PEP")

fit1 <-lm(Observed~Type, data=indice_diversite_soil)
hist(fit1$residuals) #suit une loi normale?
shapiro.test(fit1$residuals)# si p-value > 0,05 alors la distribution n'est pas significativement différente d'une distribution normale
summary(fit1)

fit2 <-lm(Shannon~Type, data=indice_diversite_soil)
hist(fit2$residuals) #suit une loi normale?
shapiro.test(fit2$residuals)
summary(fit2)

fit3 <-lm(Simpson~Type, data=indice_diversite_soil)
hist(fit3$residuals) #suit une loi normale? if not normal try a GLM
shapiro.test(fit3$residuals)
glm_fit3 <- glm(Simpson~Type, data=indice_diversite_soil) #link="inverse" by default
summary(glm_fit3)
```

Alpha-diversity in unplanted soils...
```{r}
p_Div_unplanted=plot_richness(phylo_rar_unplanted, color="Type", x="Type", measures=c("Observed", "Shannon", "Simpson"))

p_Div_unplanted_bp=p_Div_unplanted+geom_boxplot(aes(color=Type), alpha=0.2)+ theme_classic()+xlab("Alpha-diversity in unplanted samples")+ylab("Observed ASVs richness")+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))

p_Div_unplanted_bp

# en noir et blanc avec des formes
p_Div_unplanted=plot_richness(phylo_rar_unplanted, shape="Type", x="Type", measures=c("Observed", "Shannon", "Simpson"))
p_Div_unplanted_bp_nb <- p_Div_unplanted+geom_boxplot(aes(shape=Type), alpha=0.2) + scale_shape_manual(values=c(0,1,2,3,19,15,8))+ theme_classic()+xlab("Alpha-diversity in unplanted samples")+ylab("Observed ASVs richness")+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))
p_Div_unplanted_bp_nb

#Statistical analyses: be careful, if updating code: this part has to be manually (export/import)

alpha.diversity_unplanted=estimate_richness(phylo_rar_unplanted, measures=c("Observed","Shannon", "Simpson"))

#write.table(alpha.diversity_unplanted, ("indice_diversite_unplanted.txt"))

#Add a "Type" column and re-import as Text(readr)

library(readr)
indice_diversite_unplanted <-read_delim("C:/Users/hmiddleton/Dropbox/Projet_Mutants_miPEPs_miRNA/Harriet/Metabarcoding_mutants_and_miPEP/Analyses_métabarcoding/miPEP_H/16S/indice_diversite_unplanted_1.txt","\t", escape_double = FALSE, trim_ws = TRUE)

indice_diversite_unplanted$Type <- as.factor(indice_diversite_unplanted$Type)
str(object = indice_diversite_unplanted$Type)
indice_diversite_unplanted$Type <- relevel(indice_diversite_unplanted$Type, ref="Control-mi-PEP")

fit1 <-lm(Observed~Type, data=indice_diversite_unplanted)
hist(fit1$residuals) #suit une loi normale?
shapiro.test(fit1$residuals)# si p-value > 0,05 alors la distribution n'est pas significativement différente d'une distribution normale
summary(fit1)

fit2 <-lm(Shannon~Type, data=indice_diversite_unplanted)
hist(fit2$residuals) #suit une loi normale?
shapiro.test(fit2$residuals)
summary(fit2)

fit3 <-lm(Simpson~Type, data=indice_diversite_unplanted)
hist(fit3$residuals) #suit une loi normale? if not normal try a GLM
shapiro.test(fit3$residuals)
summary(fit3)
```

Bray-Curtis in root samples...
```{r}
set.seed(711)

PCOA_Bray_roots <- ordinate(physeq = phylo_roots, method = "PCoA", distance="bray") 
PCOA_Bray_roots 
# The 2 main axes: Relative_eig has to be > Broken_stick, i.e. more than just random interpretation OR at least cumul eig > cumul broken stick
otu_roots <- otu_table(phylo_roots)
rda <- rda(otu_roots) #raw data without the sample column which conflicts w/rda
screeplot(rda, bstick=TRUE)
#This graph shows that the 2nd axis is not interpretable


bc_roots <- plot_ordination(physeq = phylo_roots, ordination=PCOA_Bray_roots, color="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))+stat_ellipse()
bc_roots <- bc_roots + ggtitle("Bray-Curtis distances based PCoA of roots and close rhizospheric bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))
bc_roots

# en noir et blanc avec des formes

bc_roots_nb <- plot_ordination(physeq = phylo_roots, ordination=PCOA_Bray_roots, shape="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_shape_manual(values=c(0,1,2,3,19,15,8))+stat_ellipse()
bc_roots_nb  + ggtitle("Bray-Curtis distances based PCoA of roots and close rhizospheric bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))

#Statistical analysis: permanova & beta-disper test
# Calculate bray curtis distance matrix
bray_roots <- phyloseq::distance(phylo_roots, method = "bray")
sampledf_roots <- data.frame(sample_data(phylo_roots))
adonis(bray_roots ~ Type, data = sampledf_roots)

# If Type has a significant effect, check group dispersion
Disper_Type_roots = betadisper(bray_roots, sampledf_roots$Type)
permutest(Disper_Type_roots, pairwise=TRUE, permutations=1000)

#If the dispersion test is signif, it means that the Type effect is acutally only due to group dispersion, and not a real biological effect

#### Pairwise multilevel comparison using adonis #####
library(devtools)
#install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis") 

pairwise.adonis <- function(x,factors, sim.function = 'vegdist', sim.method = 'bray', p.adjust.m ='bonferroni',reduce=NULL,perm=999)
{
  
  co <- combn(unique(as.character(factors)),2)
  pairs <- c()
  Df <- c()
  SumsOfSqs <- c()
  F.Model <- c()
  R2 <- c()
  p.value <- c()
  
  
  for(elem in 1:ncol(co)){
    if(inherits(x, 'dist')){
      x1=as.matrix(x)[factors %in% c(as.character(co[1,elem]),as.character(co[2,elem])),
                      factors %in% c(as.character(co[1,elem]),as.character(co[2,elem]))]
    }
    
    else  (
      if (sim.function == 'daisy'){
        x1 = daisy(x[factors %in% c(co[1,elem],co[2,elem]),],metric=sim.method)
      } 
      else{x1 = vegdist(x[factors %in% c(co[1,elem],co[2,elem]),],method=sim.method)}
    )
    
    ad <- adonis(x1 ~ factors[factors %in% c(co[1,elem],co[2,elem])],
                 permutations = perm);
    pairs <- c(pairs,paste(co[1,elem],'vs',co[2,elem]));
    Df <- c(Df,ad$aov.tab[1,1])
    SumsOfSqs <- c(SumsOfSqs, ad$aov.tab[1,2])
    F.Model <- c(F.Model,ad$aov.tab[1,4]);
    R2 <- c(R2,ad$aov.tab[1,5]);
    p.value <- c(p.value,ad$aov.tab[1,6])
  }
  p.adjusted <- p.adjust(p.value,method=p.adjust.m)
  
  sig = c(rep('',length(p.adjusted)))
  sig[p.adjusted <= 0.05] <-'.'
  sig[p.adjusted <= 0.01] <-'*'
  sig[p.adjusted <= 0.001] <-'**'
  sig[p.adjusted <= 0.0001] <-'***'
  pairw.res <- data.frame(pairs,Df,SumsOfSqs,F.Model,R2,p.value,p.adjusted,sig)
  
  if(!is.null(reduce)){
    pairw.res <- subset (pairw.res, grepl(reduce,pairs))
    pairw.res$p.adjusted <- p.adjust(pairw.res$p.value,method=p.adjust.m)
    
    sig = c(rep('',length(pairw.res$p.adjusted)))
    sig[pairw.res$p.adjusted <= 0.1] <-'.'
    sig[pairw.res$p.adjusted <= 0.05] <-'*'
    sig[pairw.res$p.adjusted <= 0.01] <-'**'
    sig[pairw.res$p.adjusted <= 0.001] <-'***'
    pairw.res <- data.frame(pairw.res[,1:7],sig)
  }
  class(pairw.res) <- c("pwadonis", "data.frame")
  return(pairw.res)
} 


### Method summary
summary.pwadonis = function(object, ...) {
  cat("Result of pairwise.adonis:\n")
  cat("\n")
  print(object, ...)
  cat("\n")
  cat("Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1\n")
}

# All Types
pairwise.adonis(bray_roots, sampledf_roots$Type)
```

Bray-Curtis in soil samples...
```{r}
set.seed(711)
PCOA_Bray_soil <- ordinate(physeq = phylo_soil, method = "PCoA", distance="bray") 
PCOA_Bray_soil 
# The 2 main axes: Relative_eig has to be > Broken_stick, i.e. more than just random interpretation OR at least cumul eig > cumul broken stick
otu_soil <- otu_table(phylo_soil)
rda <- rda(otu_soil) #raw data without the sample column which conflicts w/rda
screeplot(rda, bstick=TRUE)
#This graph shows that the 2 first axes are > random

bc_soil <- plot_ordination(physeq = phylo_soil, ordination=PCOA_Bray_soil, color="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))+stat_ellipse()
bc_soil <- bc_soil + ggtitle("Bray-Curtis distances based PCoA of distant rhizospheric bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))
bc_soil

# en noir et blanc avec des formes
bc_soil_nb <- plot_ordination(physeq = phylo_soil, ordination=PCOA_Bray_soil, shape="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_shape_manual(values=c(0,1,2,3,19,15,8))+stat_ellipse()
bc_soil_nb <- bc_soil_nb  + ggtitle("Bray-Curtis distances based PCoA of distant rhizospheric bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))
bc_soil_nb

#Statistical analysis: permanova & beta-disper test
# Calculate bray curtis distance matrix
bray_soil <- phyloseq::distance(phylo_soil, method = "bray")
sampledf_soil <- data.frame(sample_data(phylo_soil))
adonis(bray_soil ~ Type, data = sampledf_soil)

# If Type has a significant effect, check group dispersion
Disper_Type_soil = betadisper(bray_soil, sampledf_soil$Type)
permutest(Disper_Type_soil, pairwise=TRUE, permutations=1000)

#If the dispersion test is signif, it means that the Type effect is acutally only due to group dispersion, and not a real biological effect

pairwise.adonis(bray_soil, sampledf_soil$Type)
```

Bray-Curtis in unplanted samples...
```{r}
set.seed(711)
PCOA_Bray_unplanted <- ordinate(physeq = phylo_unplanted, method = "PCoA", distance="bray") 
PCOA_Bray_unplanted 
# The 2 main axes: Relative_eig has to be > Broken_stick, i.e. more than just random interpretation OR at least cumul eig > cumul broken stick
otu_unplanted <- otu_table(phylo_unplanted)
rda <- rda(otu_unplanted) #raw data without the sample column which conflicts w/rda
screeplot(rda, bstick=TRUE)
#This graph shows that the 2nd axis is not interpretable


bc_unplanted <- plot_ordination(physeq = phylo_unplanted, ordination=PCOA_Bray_unplanted, color="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))+stat_ellipse()
bc_unplanted <- bc_unplanted + ggtitle("Bray-Curtis distances based PCoA of unplanted soil bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))
bc_unplanted

# en noir et blanc avec des formes
bc_unplanted_nb <- plot_ordination(physeq = phylo_unplanted, ordination=PCOA_Bray_unplanted, shape="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_shape_manual(values=c(0,1,2,3,19,15,8))+stat_ellipse()
bc_unplanted_nb <- bc_unplanted_nb  + ggtitle("Bray-Curtis distances based PCoA of unplanted soil bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))
bc_unplanted_nb

#Statistical analysis: permanova & beta-disper test
# Calculate bray curtis distance matrix
bray_unplanted <- phyloseq::distance(phylo_unplanted, method = "bray")
sampledf_unplanted <- data.frame(sample_data(phylo_unplanted))
adonis(bray_unplanted ~ Type, data = sampledf_unplanted)

# If Type has a significant effect, check group dispersion
Disper_Type_unplanted = betadisper(bray_unplanted, sampledf_unplanted$Type)
permutest(Disper_Type_unplanted, pairwise=TRUE, permutations=1000)

#If the dispersion test is signif, it means that the Type effect is acutally only due to group dispersion, and not a real biological effect

pairwise.adonis(bray_unplanted, sampledf_unplanted$Type)
```


Weighted UniFrac : first, add phylogenetic tree
```{r}
library("ape")
library(magrittr)

pick_new_outgroup <- function(tree.unrooted){
  require("magrittr")
  require("data.table")
  require("ape") # ape::Ntip
  # tablify parts of tree that we need.
  treeDT <-
    cbind(
      data.table(tree.unrooted$edge),
      data.table(length = tree.unrooted$edge.length)
    )[1:Ntip(tree.unrooted)] %>%
    cbind(data.table(id = tree.unrooted$tip.label))
  # Take the longest terminal branch as outgroup
  new.outgroup <- treeDT[which.max(length)]$id
  return(new.outgroup) }
  
tree <- read.tree("~/metabarcoding_miPEP_mutants/16S/16S/plaque2/export/tree/tree.fasttree")
tree
my.tree <- phy_tree(tree)
out.group <- pick_new_outgroup(my.tree)
out.group #ex : [1] "3312"
new.tree <- ape::root(my.tree, outgroup=out.group, resolve.root=TRUE)#root tree to always have the same PCoA at each re-run
new.tree
#use the latest phyloseq info 
phylo_for_UniFrac <- phyloseq(tax_table,OTU_table,samples, new.tree)
phylo_for_UniFrac
```

Subsample the UniFrac phyloseq object
```{r}
correct.order <- c("miPEP-A","miPEP-B","miPEP-C","Control-mi-PEP")
sample_data(phylo_for_UniFrac)$Type <- factor(sample_data(phylo_for_UniFrac)$Type, levels = correct.order)
levels(get_variable(phylo_for_UniFrac, "Type")) #mettre dans le bon ordre

phylo_for_UniFrac_soil <- subset_samples(phylo_for_UniFrac, Compartment=="SOL")
phylo_for_UniFrac_roots <- subset_samples(phylo_for_UniFrac, Compartment=="RACINE")
phylo_for_UniFrac_unplanted <- subset_samples(phylo_for_UniFrac, Compartment=="SOL.NU")
```

Weighted UniFrac in root samples...
```{r}
set.seed(711)
PCOA_weighted_roots <- ordinate(physeq = phylo_for_UniFrac_roots,  method = "PCoA", distance="unifrac", weighted=TRUE) 
PCOA_weighted_roots 
# The 2 main axes: Relative_eig has to be > Broken_stick, i.e. more than just random interpretation OR at least cumul eig > cumul broken stick
otu_w_roots <- otu_table(phylo_for_UniFrac_roots)
rda <- rda(otu_w_roots) #raw data without the sample column which conflicts w/rda
screeplot(rda, bstick=TRUE)
#This graph shows that the 2nd axis is random

WUF_roots <- plot_ordination(physeq = phylo_for_UniFrac_roots, ordination=PCOA_weighted_roots, color="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))+stat_ellipse()
WUF_roots <- WUF_roots + ggtitle("Weighted UniFrac distances based PCoA of roots and close rhizospheric bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))
WUF_roots

# en noir et blanc avec des formes
WUF_roots_nb <- plot_ordination(physeq = phylo_for_UniFrac_roots, ordination=PCOA_weighted_roots, shape="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_shape_manual(values=c(0,1,2,3,19,15,8))+stat_ellipse()
WUF_roots_nb <- WUF_roots_nb  + ggtitle("Weighted UniFrac distances based PCoA of roots and close rhizospheric bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))

#Statistical analysis: permanova & beta-disper test

# Calculate distance matrix
WUF_roots <- phyloseq::distance(phylo_for_UniFrac_roots, method = "unifrac", weighted=TRUE)
sampledf_roots <- data.frame(sample_data(phylo_for_UniFrac_roots))
adonis(WUF_roots ~ Type, data = sampledf_roots)

# If Type has a significant effect, check group dispersion
Disper_Type_roots = betadisper(WUF_roots, sampledf_roots$Type)
permutest(Disper_Type_roots, pairwise=TRUE, permutations=1000)

#If the dispersion test is signif, it means that the Type effect is acutally only due to group dispersion, and not a real biological effect

pairwise.adonis(WUF_roots, sampledf_roots$Type)
```

Weighted UniFrac in soil samples...
```{r}
set.seed(711)
PCOA_weighted_soil <- ordinate(physeq = phylo_for_UniFrac_soil,  method = "PCoA", distance="unifrac", weighted=TRUE) 
PCOA_weighted_roots 
# The 2 main axes: Relative_eig has to be > Broken_stick, i.e. more than just random interpretation OR at least cumul eig > cumul broken stick
otu_w_roots <- otu_table(phylo_for_UniFrac_roots)
rda <- rda(otu_w_roots) #raw data without the sample column which conflicts w/rda
screeplot(rda, bstick=TRUE)
#This graph shows that the 2nd axis is random

WUF_soil <- plot_ordination(physeq = phylo_for_UniFrac_soil, ordination=PCOA_weighted_soil, color="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))+stat_ellipse()
WUF_soil <- WUF_soil + ggtitle("Weighted UniFrac distances based PCoA of distant rhizospheric bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))
WUF_soil

# en noir et blanc avec des formes
WUF_soil_nb <- plot_ordination(physeq = phylo_for_UniFrac_soil, ordination=PCOA_weighted_soil, shape="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_shape_manual(values=c(0,1,2,3,19,15,8))+stat_ellipse()
WUF_soil_nb <- WUF_soil_nb  + ggtitle("Weighted UniFrac distances based PCoA of distant rhizospheric bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))

#Statistical analysis: permanova & beta-disper test

# Calculate distance matrix
WUF_soil <- phyloseq::distance(phylo_for_UniFrac_soil, method = "unifrac", weighted=TRUE)
sampledf_soil <- data.frame(sample_data(phylo_for_UniFrac_soil))
adonis(WUF_soil ~ Type, data = sampledf_soil)

# If Type has a significant effect, check group dispersion
Disper_Type_soil = betadisper(WUF_soil, sampledf_soil$Type)
permutest(Disper_Type_soil, pairwise=TRUE, permutations=1000)

#If the dispersion test is signif, it means that the Type effect is acutally only due to group dispersion, and not a real biological effect
pairwise.adonis(WUF_soil, sampledf_soil$Type)
```

Weighted UniFrac in unplanted samples...
```{r}
set.seed(711)
PCOA_weighted_unplanted <- ordinate(physeq = phylo_for_UniFrac_unplanted,  method = "PCoA", distance="unifrac", weighted=TRUE) 
PCOA_weighted_unplanted 
# The 2 main axes: Relative_eig has to be > Broken_stick, i.e. more than just random interpretation OR at least cumul eig > cumul broken stick
otu_w_unplanted <- otu_table(phylo_for_UniFrac_unplanted)
rda <- rda(otu_w_unplanted) #raw data without the sample column which conflicts w/rda
screeplot(rda, bstick=TRUE)
#This graph shows that the 2nd axis is random

WUF_unplanted <- plot_ordination(physeq = phylo_for_UniFrac_unplanted, ordination=PCOA_weighted_unplanted, color="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))+stat_ellipse()
WUF_unplanted <- WUF_unplanted + ggtitle("Weighted UniFrac distances based PCoA of unplanted soil bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))
WUF_unplanted

# en noir et blanc avec des formes
WUF_unplanted_nb <- plot_ordination(physeq = phylo_for_UniFrac_unplanted, ordination=PCOA_weighted_unplanted, shape="Type")+ theme_classic()+ theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+scale_shape_manual(values=c(0,1,2,3,19,15,8))+stat_ellipse()
WUF_unplanted_nb <- WUF_unplanted_nb  + ggtitle("Weighted UniFrac distances based PCoA of unplanted soil bacterial communities") +theme(plot.title = element_text(size = 8, face = "bold"))

#Statistical analysis: permanova & beta-disper test

# Calculate distance matrix
WUF_unplanted <- phyloseq::distance(phylo_for_UniFrac_unplanted, method = "unifrac", weighted=TRUE)
sampledf_unplanted <- data.frame(sample_data(phylo_for_UniFrac_unplanted))
adonis(WUF_unplanted ~ Type, data = sampledf_unplanted)

# If Type has a significant effect, check group dispersion
Disper_Type_unplanted = betadisper(WUF_unplanted, sampledf_unplanted$Type)
permutest(Disper_Type_unplanted, pairwise=TRUE, permutations=1000)

#If the dispersion test is signif, it means that the Type effect is acutally only due to group dispersion, and not a real biological effect

pairwise.adonis(WUF_unplanted, sampledf_unplanted$Type)
```

Another type of analysis is to look at specific OTUs/ASVs that are specifically enriched or depleted in certain sample Types 

1) Root samples
```{r}
library(DESeq2) # The DESeq2 package uses the raw OTU/ASV table (non rarefied phyloseq) and calculates a negative binomial GLM for each OTU/ASV , the Wald test for stats

# Subset phylo_roots to have only the control and miPEP-A

phylo_roots_1 <- subset_samples(phylo_roots, Type=="Control-mi-PEP")
phylo_roots_2 <- subset_samples(phylo_roots, Type=="miPEP-A")
phylo_roots_A <- merge_phyloseq(phylo_roots_1,phylo_roots_2)

rm(phylo_roots_1, phylo_roots_2)

#Convert phyloseq object into a DESeqDataSet object with dispersion estimates using "Type" as the main factor
ds2_roots <- phyloseq_to_deseq2(phylo_roots_A, ~ Type) 
#Specify your reference group, or else it will use alphabetical order
ds2_roots$Type <- relevel(ds2_roots$Type, ref = "Control-mi-PEP")  

dds_roots <- DESeq(ds2_roots, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_roots) >= 10 ) >=3 #filters out ASVs present in less than 3 samples in this subset (under 10 reads)
dds_roots <- dds_roots[keep,]

resultsNames(dds_roots) #prints the names of the coefficients

##miPEP-A##
resLFC <- lfcShrink(dds_roots, coef="Type_miPEP.A_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_roots_miPEPa_order = resLFC[order(resLFC$padj, na.last=NA), ] #570ASVs
sigtab_roots_miPEPa = res_roots_miPEPa_order[which(res_roots_miPEPa_order$padj < alpha),] #55 ASVs
sigtab_roots_miPEPa_top_50 <- sigtab_roots_miPEPa[c(1:50),] 
sigtab_roots_miPEPa_top_50 = cbind(as(sigtab_roots_miPEPa_top_50, "data.frame"), as(tax_table(phylo_roots)[rownames(sigtab_roots_miPEPa_top_50), ], "matrix"))
dim(sigtab_roots_miPEPa_top_50)

########### Graphical representation #####
library("ggplot2")
theme_set(theme_bw())
scale_fill_discrete <- function(palname = "Set1", ...) {
    scale_fill_brewer(palette = palname, ...)
}
# Phylum order
x = tapply(sigtab_roots_miPEPa_top_50$log2FoldChange, sigtab_roots_miPEPa_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_roots_miPEPa_top_50$Phylum = factor(as.character(sigtab_roots_miPEPa_top_50$Phylum), levels=names(x))

sigtab_roots_miPEPa_top_50 <- tibble::rownames_to_column(sigtab_roots_miPEPa_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_roots_miPEPa_top_50$log2FoldChange, sigtab_roots_miPEPa_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_roots_miPEPa_top_50$VALUE = factor(as.character(sigtab_roots_miPEPa_top_50$VALUE), levels=names(x))
plot_miPEPa_roots <- ggplot(sigtab_roots_miPEPa_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-A in roots and close rhizosphere")

sigtab_roots_miPEPa_order <- order(sigtab_roots_miPEPa_top_50$log2FoldChange)
sigtab_roots_miPEPa_order = sigtab_roots_miPEPa_top_50[order(sigtab_roots_miPEPa_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_roots_miPEPa_order$Genus)

plot_miPEPa_roots_order <- plot_miPEPa_roots + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c( "darkorchid", "darkolivegreen1","brown1","firebrick", "lightskyblue","darkgreen", "khaki2","deeppink","darksalmon"))
plot_miPEPa_roots_order <- plot_miPEPa_roots_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPa_roots_order

rm(phylo_roots_A)

##miPEP-B##

phylo_roots_1 <- subset_samples(phylo_roots, Type=="Control-mi-PEP")
phylo_roots_2 <- subset_samples(phylo_roots, Type=="miPEP-B")
phylo_roots_B <- merge_phyloseq(phylo_roots_1,phylo_roots_2)

rm(phylo_roots_1, phylo_roots_2)

ds2_roots <- phyloseq_to_deseq2(phylo_roots_B, ~ Type) 

ds2_roots$Type <- relevel(ds2_roots$Type, ref = "Control-mi-PEP")  

dds_roots <- DESeq(ds2_roots, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_roots) >= 10 ) >=3
dds_roots <- dds_roots[keep,]

resLFC <- lfcShrink(dds_roots, coef="Type_miPEP.B_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_roots_miPEPb_order = resLFC[order(resLFC$padj, na.last=NA), ] #551 ASVs
sigtab_roots_miPEPb = res_roots_miPEPb_order[which(res_roots_miPEPb_order$padj < alpha),] #69 ASVs
sigtab_roots_miPEPb_top_50 <- sigtab_roots_miPEPb[c(1:50),] 
sigtab_roots_miPEPb_top_50 = cbind(as(sigtab_roots_miPEPb_top_50, "data.frame"), as(tax_table(phylo_roots)[rownames(sigtab_roots_miPEPb_top_50), ], "matrix"))
dim(sigtab_roots_miPEPb_top_50)


########### Graphical representation #####

# Phylum order
x = tapply(sigtab_roots_miPEPb_top_50$log2FoldChange, sigtab_roots_miPEPb_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_roots_miPEPb_top_50$Phylum = factor(as.character(sigtab_roots_miPEPb_top_50$Phylum), levels=names(x))

sigtab_roots_miPEPb_top_50 <- tibble::rownames_to_column(sigtab_roots_miPEPb_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_roots_miPEPb_top_50$log2FoldChange, sigtab_roots_miPEPb_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_roots_miPEPb_top_50$VALUE = factor(as.character(sigtab_roots_miPEPb_top_50$VALUE), levels=names(x))
plot_miPEPb_roots <- ggplot(sigtab_roots_miPEPb_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-B in roots and close rhizosphere")

sigtab_roots_miPEPb_order = sigtab_roots_miPEPb_top_50[order(sigtab_roots_miPEPb_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_roots_miPEPb_order$Genus)

plot_miPEPb_roots_order <- plot_miPEPb_roots + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c("darkorchid", "darkolivegreen1","brown1", "deeppink","darkblue",   "firebrick",  "darksalmon","darkcyan","darkorange" ))
plot_miPEPb_roots_order <- plot_miPEPb_roots_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPb_roots_order

rm(phylo_roots_B)

##miPEP-C##
phylo_roots_1 <- subset_samples(phylo_roots, Type=="Control-mi-PEP")
phylo_roots_2 <- subset_samples(phylo_roots, Type=="miPEP-C")
phylo_roots_C <- merge_phyloseq(phylo_roots_1,phylo_roots_2)

rm(phylo_roots_1, phylo_roots_2)

ds2_roots <- phyloseq_to_deseq2(phylo_roots_C, ~ Type) 

ds2_roots$Type <- relevel(ds2_roots$Type, ref = "Control-mi-PEP")  

dds_roots <- DESeq(ds2_roots, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_roots) >= 10 ) >=3
dds_roots <- dds_roots[keep,]

resLFC <- lfcShrink(dds_roots, coef="Type_miPEP.C_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_roots_miPEPc_order = resLFC[order(resLFC$padj, na.last=NA), ] #584
sigtab_roots_miPEPc = res_roots_miPEPc_order[which(res_roots_miPEPc_order$padj < alpha),] #59
sigtab_roots_miPEPc_top_50 <- sigtab_roots_miPEPc[c(1:50),] 
sigtab_roots_miPEPc_top_50 = cbind(as(sigtab_roots_miPEPc_top_50, "data.frame"), as(tax_table(phylo_roots)[rownames(sigtab_roots_miPEPc_top_50), ], "matrix"))
dim(sigtab_roots_miPEPc_top_50)


########### Graphical representation #####

# Phylum order
x = tapply(sigtab_roots_miPEPc_top_50$log2FoldChange, sigtab_roots_miPEPc_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_roots_miPEPc_top_50$Phylum = factor(as.character(sigtab_roots_miPEPc_top_50$Phylum), levels=names(x))

sigtab_roots_miPEPc_top_50 <- tibble::rownames_to_column(sigtab_roots_miPEPc_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_roots_miPEPc_top_50$log2FoldChange, sigtab_roots_miPEPc_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_roots_miPEPc_top_50$VALUE = factor(as.character(sigtab_roots_miPEPc_top_50$VALUE), levels=names(x))
plot_miPEPc_roots <- ggplot(sigtab_roots_miPEPc_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-C in roots and close rhizosphere")

sigtab_roots_miPEPc_order = sigtab_roots_miPEPc_top_50[order(sigtab_roots_miPEPc_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_roots_miPEPc_order$Genus)

plot_miPEPc_roots_order <- plot_miPEPc_roots + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c("darkolivegreen1","darkorchid", "darkblue","brown1","deeppink", "firebrick","darkcyan","darksalmon"))
plot_miPEPc_roots_order <- plot_miPEPc_roots_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPc_roots_order

rm(phylo_roots_C)
```

2) Soil samples
```{r}
library(DESeq2) # The DESeq2 package uses the raw OTU/ASV table (non rarefied phyloseq) and calculates a negative binomial GLM for each OTU/ASV , the Wald test for stats

# Subset phylo_soil to have only the control and miPEP-A

phylo_soil_1 <- subset_samples(phylo_soil, Type=="Control-mi-PEP")
phylo_soil_2 <- subset_samples(phylo_soil, Type=="miPEP-A")
phylo_soil_A <- merge_phyloseq(phylo_soil_1,phylo_soil_2)

rm(phylo_soil_1, phylo_soil_2)

#Convert phyloseq object into a DESeqDataSet object with dispersion estimates using "Type" as the main factor
ds2_soil <- phyloseq_to_deseq2(phylo_soil_A, ~ Type) 
#Specify your reference group, or else it will use alphabetical order
ds2_soil$Type <- relevel(ds2_soil$Type, ref = "Control-mi-PEP")  

dds_soil <- DESeq(ds2_soil, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_soil) >= 10 ) >=3 #filters out ASVs present in less than 3 samples in this subset (under 10 reads)
dds_soil <- dds_soil[keep,]

resultsNames(dds_soil) #prints the names of the coefficients

##miPEP-A##
resLFC <- lfcShrink(dds_soil, coef="Type_miPEP.A_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_soil_miPEPa_order = resLFC[order(resLFC$padj, na.last=NA), ] #834ASVs
sigtab_soil_miPEPa = res_soil_miPEPa_order[which(res_soil_miPEPa_order$padj < alpha),] #106 ASVs
sigtab_soil_miPEPa_top_50 <- sigtab_soil_miPEPa[c(1:50),] 
sigtab_soil_miPEPa_top_50 = cbind(as(sigtab_soil_miPEPa_top_50, "data.frame"), as(tax_table(phylo_soil)[rownames(sigtab_soil_miPEPa_top_50), ], "matrix"))
dim(sigtab_soil_miPEPa_top_50)

########### Graphical representation #####
library("ggplot2")
theme_set(theme_bw())
scale_fill_discrete <- function(palname = "Set1", ...) {
    scale_fill_brewer(palette = palname, ...)
}
# Phylum order
x = tapply(sigtab_soil_miPEPa_top_50$log2FoldChange, sigtab_soil_miPEPa_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_soil_miPEPa_top_50$Phylum = factor(as.character(sigtab_soil_miPEPa_top_50$Phylum), levels=names(x))

sigtab_soil_miPEPa_top_50 <- tibble::rownames_to_column(sigtab_soil_miPEPa_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_soil_miPEPa_top_50$log2FoldChange, sigtab_soil_miPEPa_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_soil_miPEPa_top_50$VALUE = factor(as.character(sigtab_soil_miPEPa_top_50$VALUE), levels=names(x))
plot_miPEPa_soil <- ggplot(sigtab_soil_miPEPa_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-A in distant rhizosphere")

sigtab_soil_miPEPa_order <- order(sigtab_soil_miPEPa_top_50$log2FoldChange)
sigtab_soil_miPEPa_order = sigtab_soil_miPEPa_top_50[order(sigtab_soil_miPEPa_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_soil_miPEPa_order$Genus)

plot_miPEPa_soil_order <- plot_miPEPa_soil + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c( "darkorchid", "darkolivegreen1","brown1","firebrick", "lightskyblue","darkgreen", "khaki2","deeppink","darksalmon"))
plot_miPEPa_soil_order <- plot_miPEPa_soil_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPa_soil_order

rm(phylo_soil_A)

##miPEP-B##

phylo_soil_1 <- subset_samples(phylo_soil, Type=="Control-mi-PEP")
phylo_soil_2 <- subset_samples(phylo_soil, Type=="miPEP-B")
phylo_soil_B <- merge_phyloseq(phylo_soil_1,phylo_soil_2)

rm(phylo_soil_1, phylo_soil_2)

ds2_soil <- phyloseq_to_deseq2(phylo_soil_B, ~ Type) 

ds2_soil$Type <- relevel(ds2_soil$Type, ref = "Control-mi-PEP")  

dds_soil <- DESeq(ds2_soil, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_soil) >= 10 ) >=3
dds_soil <- dds_soil[keep,]

resLFC <- lfcShrink(dds_soil, coef="Type_miPEP.B_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_soil_miPEPb_order = resLFC[order(resLFC$padj, na.last=NA), ] #985 ASVs
sigtab_soil_miPEPb = res_soil_miPEPb_order[which(res_soil_miPEPb_order$padj < alpha),] #100 ASVs
sigtab_soil_miPEPb_top_50 <- sigtab_soil_miPEPb[c(1:50),] 
sigtab_soil_miPEPb_top_50 = cbind(as(sigtab_soil_miPEPb_top_50, "data.frame"), as(tax_table(phylo_soil)[rownames(sigtab_soil_miPEPb_top_50), ], "matrix"))
dim(sigtab_soil_miPEPb_top_50)


########### Graphical representation #####

# Phylum order
x = tapply(sigtab_soil_miPEPb_top_50$log2FoldChange, sigtab_soil_miPEPb_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_soil_miPEPb_top_50$Phylum = factor(as.character(sigtab_soil_miPEPb_top_50$Phylum), levels=names(x))

sigtab_soil_miPEPb_top_50 <- tibble::rownames_to_column(sigtab_soil_miPEPb_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_soil_miPEPb_top_50$log2FoldChange, sigtab_soil_miPEPb_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_soil_miPEPb_top_50$VALUE = factor(as.character(sigtab_soil_miPEPb_top_50$VALUE), levels=names(x))
plot_miPEPb_soil <- ggplot(sigtab_soil_miPEPb_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-B in distant rhizosphere")

sigtab_soil_miPEPb_order = sigtab_soil_miPEPb_top_50[order(sigtab_soil_miPEPb_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_soil_miPEPb_order$Genus)

plot_miPEPb_soil_order <- plot_miPEPb_soil + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c("darkorchid", "darkolivegreen1","brown1", "deeppink","darkblue",   "firebrick",  "darksalmon","darkcyan","darkorange" ))
plot_miPEPb_soil_order <- plot_miPEPb_soil_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPb_soil_order

rm(phylo_soil_B)

##miPEP-C##
phylo_soil_1 <- subset_samples(phylo_soil, Type=="Control-mi-PEP")
phylo_soil_2 <- subset_samples(phylo_soil, Type=="miPEP-C")
phylo_soil_C <- merge_phyloseq(phylo_soil_1,phylo_soil_2)

rm(phylo_soil_1, phylo_soil_2)

ds2_soil <- phyloseq_to_deseq2(phylo_soil_C, ~ Type) 

ds2_soil$Type <- relevel(ds2_soil$Type, ref = "Control-mi-PEP")  

dds_soil <- DESeq(ds2_soil, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_soil) >= 10 ) >=3
dds_soil <- dds_soil[keep,]

resLFC <- lfcShrink(dds_soil, coef="Type_miPEP.C_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_soil_miPEPc_order = resLFC[order(resLFC$padj, na.last=NA), ] #937
sigtab_soil_miPEPc = res_soil_miPEPc_order[which(res_soil_miPEPc_order$padj < alpha),] #106
sigtab_soil_miPEPc_top_50 <- sigtab_soil_miPEPc[c(1:50),] 
sigtab_soil_miPEPc_top_50 = cbind(as(sigtab_soil_miPEPc_top_50, "data.frame"), as(tax_table(phylo_soil)[rownames(sigtab_soil_miPEPc_top_50), ], "matrix"))
dim(sigtab_soil_miPEPc_top_50)


########### Graphical representation #####

# Phylum order
x = tapply(sigtab_soil_miPEPc_top_50$log2FoldChange, sigtab_soil_miPEPc_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_soil_miPEPc_top_50$Phylum = factor(as.character(sigtab_soil_miPEPc_top_50$Phylum), levels=names(x))

sigtab_soil_miPEPc_top_50 <- tibble::rownames_to_column(sigtab_soil_miPEPc_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_soil_miPEPc_top_50$log2FoldChange, sigtab_soil_miPEPc_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_soil_miPEPc_top_50$VALUE = factor(as.character(sigtab_soil_miPEPc_top_50$VALUE), levels=names(x))
plot_miPEPc_soil <- ggplot(sigtab_soil_miPEPc_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-C in distant rhizosphere")

sigtab_soil_miPEPc_order = sigtab_soil_miPEPc_top_50[order(sigtab_soil_miPEPc_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_soil_miPEPc_order$Genus)

plot_miPEPc_soil_order <- plot_miPEPc_soil + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c("darkolivegreen1","darkorchid", "darkblue","brown1","deeppink", "firebrick","darkcyan","darksalmon", "green"))
plot_miPEPc_soil_order <- plot_miPEPc_soil_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPc_soil_order

rm(phylo_soil_C)
```

3) Unplanted samples
```{r}
library(DESeq2) # The DESeq2 package uses the raw OTU/ASV table (non rarefied phyloseq) and calculates a negative binomial GLM for each OTU/ASV , the Wald test for stats

# Subset phylo_unplanted to have only the control and miPEP-A

phylo_unplanted_1 <- subset_samples(phylo_unplanted, Type=="Control-mi-PEP")
phylo_unplanted_2 <- subset_samples(phylo_unplanted, Type=="miPEP-A")
phylo_unplanted_A <- merge_phyloseq(phylo_unplanted_1,phylo_unplanted_2)

rm(phylo_unplanted_1, phylo_unplanted_2)

#Convert phyloseq object into a DESeqDataSet object with dispersion estimates using "Type" as the main factor
ds2_unplanted <- phyloseq_to_deseq2(phylo_unplanted_A, ~ Type) 
#Specify your reference group, or else it will use alphabetical order
ds2_unplanted$Type <- relevel(ds2_unplanted$Type, ref = "Control-mi-PEP")  

dds_unplanted <- DESeq(ds2_unplanted, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_unplanted) >= 10 ) >=3 #filters out ASVs present in less than 3 samples in this subset (under 10 reads)
dds_unplanted <- dds_unplanted[keep,]

resultsNames(dds_unplanted) #prints the names of the coefficients

##miPEP-A##
resLFC <- lfcShrink(dds_unplanted, coef="Type_miPEP.A_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_unplanted_miPEPa_order = resLFC[order(resLFC$padj, na.last=NA), ] #720
sigtab_unplanted_miPEPa = res_unplanted_miPEPa_order[which(res_unplanted_miPEPa_order$padj < alpha),] #77 ASVs
sigtab_unplanted_miPEPa_top_50 <- sigtab_unplanted_miPEPa[c(1:50),] 
sigtab_unplanted_miPEPa_top_50 = cbind(as(sigtab_unplanted_miPEPa_top_50, "data.frame"), as(tax_table(phylo_unplanted)[rownames(sigtab_unplanted_miPEPa_top_50), ], "matrix"))
dim(sigtab_unplanted_miPEPa_top_50)

########### Graphical representation #####
library("ggplot2")
theme_set(theme_bw())
scale_fill_discrete <- function(palname = "Set1", ...) {
    scale_fill_brewer(palette = palname, ...)
}
# Phylum order
x = tapply(sigtab_unplanted_miPEPa_top_50$log2FoldChange, sigtab_unplanted_miPEPa_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_unplanted_miPEPa_top_50$Phylum = factor(as.character(sigtab_unplanted_miPEPa_top_50$Phylum), levels=names(x))

sigtab_unplanted_miPEPa_top_50 <- tibble::rownames_to_column(sigtab_unplanted_miPEPa_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_unplanted_miPEPa_top_50$log2FoldChange, sigtab_unplanted_miPEPa_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_unplanted_miPEPa_top_50$VALUE = factor(as.character(sigtab_unplanted_miPEPa_top_50$VALUE), levels=names(x))
plot_miPEPa_unplanted <- ggplot(sigtab_unplanted_miPEPa_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-A in unplanted soils")

sigtab_unplanted_miPEPa_order <- order(sigtab_unplanted_miPEPa_top_50$log2FoldChange)
sigtab_unplanted_miPEPa_order = sigtab_unplanted_miPEPa_top_50[order(sigtab_unplanted_miPEPa_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_unplanted_miPEPa_order$Genus)

plot_miPEPa_unplanted_order <- plot_miPEPa_unplanted + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c( "darkorchid", "darkolivegreen1","brown1","firebrick", "lightskyblue","darkgreen", "khaki2","deeppink","darksalmon"))
plot_miPEPa_unplanted_order <- plot_miPEPa_unplanted_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPa_unplanted_order

rm(phylo_unplanted_A)

##miPEP-B##

phylo_unplanted_1 <- subset_samples(phylo_unplanted, Type=="Control-mi-PEP")
phylo_unplanted_2 <- subset_samples(phylo_unplanted, Type=="miPEP-B")
phylo_unplanted_B <- merge_phyloseq(phylo_unplanted_1,phylo_unplanted_2)

rm(phylo_unplanted_1, phylo_unplanted_2)

ds2_unplanted <- phyloseq_to_deseq2(phylo_unplanted_B, ~ Type) 

ds2_unplanted$Type <- relevel(ds2_unplanted$Type, ref = "Control-mi-PEP")  

dds_unplanted <- DESeq(ds2_unplanted, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_unplanted) >= 10 ) >=3
dds_unplanted <- dds_unplanted[keep,]

resLFC <- lfcShrink(dds_unplanted, coef="Type_miPEP.B_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_unplanted_miPEPb_order = resLFC[order(resLFC$padj, na.last=NA), ] #902 ASVs
sigtab_unplanted_miPEPb = res_unplanted_miPEPb_order[which(res_unplanted_miPEPb_order$padj < alpha),] #83 ASVs
sigtab_unplanted_miPEPb_top_50 <- sigtab_unplanted_miPEPb[c(1:50),] 
sigtab_unplanted_miPEPb_top_50 = cbind(as(sigtab_unplanted_miPEPb_top_50, "data.frame"), as(tax_table(phylo_unplanted)[rownames(sigtab_unplanted_miPEPb_top_50), ], "matrix"))
dim(sigtab_unplanted_miPEPb_top_50)


########### Graphical representation #####

# Phylum order
x = tapply(sigtab_unplanted_miPEPb_top_50$log2FoldChange, sigtab_unplanted_miPEPb_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_unplanted_miPEPb_top_50$Phylum = factor(as.character(sigtab_unplanted_miPEPb_top_50$Phylum), levels=names(x))

sigtab_unplanted_miPEPb_top_50 <- tibble::rownames_to_column(sigtab_unplanted_miPEPb_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_unplanted_miPEPb_top_50$log2FoldChange, sigtab_unplanted_miPEPb_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_unplanted_miPEPb_top_50$VALUE = factor(as.character(sigtab_unplanted_miPEPb_top_50$VALUE), levels=names(x))
plot_miPEPb_unplanted <- ggplot(sigtab_unplanted_miPEPb_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-B in unplanted soils")

sigtab_unplanted_miPEPb_order = sigtab_unplanted_miPEPb_top_50[order(sigtab_unplanted_miPEPb_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_unplanted_miPEPb_order$Genus)

plot_miPEPb_unplanted_order <- plot_miPEPb_unplanted + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c("darkorchid", "darkolivegreen1","brown1", "deeppink","darkblue",   "firebrick",  "darksalmon","darkcyan","darkorange","turquoise","magenta","green" ))
plot_miPEPb_unplanted_order <- plot_miPEPb_unplanted_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPb_unplanted_order

rm(phylo_unplanted_B)

##miPEP-C##
phylo_unplanted_1 <- subset_samples(phylo_unplanted, Type=="Control-mi-PEP")
phylo_unplanted_2 <- subset_samples(phylo_unplanted, Type=="miPEP-C")
phylo_unplanted_C <- merge_phyloseq(phylo_unplanted_1,phylo_unplanted_2)

rm(phylo_unplanted_1, phylo_unplanted_2)

ds2_unplanted <- phyloseq_to_deseq2(phylo_unplanted_C, ~ Type) 

ds2_unplanted$Type <- relevel(ds2_unplanted$Type, ref = "Control-mi-PEP")  

dds_unplanted <- DESeq(ds2_unplanted, test="Wald", fitType="parametric")

keep <- rowSums(counts(dds_unplanted) >= 10 ) >=3
dds_unplanted <- dds_unplanted[keep,]

resLFC <- lfcShrink(dds_unplanted, coef="Type_miPEP.C_vs_Control.mi.PEP", type="ashr")
resLFC

alpha=0.001

res_unplanted_miPEPc_order = resLFC[order(resLFC$padj, na.last=NA), ] #823
sigtab_unplanted_miPEPc = res_unplanted_miPEPc_order[which(res_unplanted_miPEPc_order$padj < alpha),] #78
sigtab_unplanted_miPEPc_top_50 <- sigtab_unplanted_miPEPc[c(1:50),] 
sigtab_unplanted_miPEPc_top_50 = cbind(as(sigtab_unplanted_miPEPc_top_50, "data.frame"), as(tax_table(phylo_unplanted)[rownames(sigtab_unplanted_miPEPc_top_50), ], "matrix"))
dim(sigtab_unplanted_miPEPc_top_50)


########### Graphical representation #####

# Phylum order
x = tapply(sigtab_unplanted_miPEPc_top_50$log2FoldChange, sigtab_unplanted_miPEPc_top_50$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_unplanted_miPEPc_top_50$Phylum = factor(as.character(sigtab_unplanted_miPEPc_top_50$Phylum), levels=names(x))

sigtab_unplanted_miPEPc_top_50 <- tibble::rownames_to_column(sigtab_unplanted_miPEPc_top_50, "VALUE")

# VALUE order
x = tapply(sigtab_unplanted_miPEPc_top_50$log2FoldChange, sigtab_unplanted_miPEPc_top_50$VALUE, function(x) max(x))
x = sort(x, TRUE)
sigtab_unplanted_miPEPc_top_50$VALUE = factor(as.character(sigtab_unplanted_miPEPc_top_50$VALUE), levels=names(x))
plot_miPEPc_unplanted <- ggplot(sigtab_unplanted_miPEPc_top_50, aes(x=VALUE, y=log2FoldChange, color=Phylum)) + geom_point(size=2) +theme(axis.text.x = element_text(size=8, angle = -90, hjust = 0, vjust=0.5), legend.text = element_text(size = 10))+ geom_hline(yintercept=0, color="black") + labs(title = "Top 50 differentially abundant ASVs in comparison with control (p.adj<0.001)",subtitle = "MiPEP-C in unplanted soils")

sigtab_unplanted_miPEPc_order = sigtab_unplanted_miPEPc_top_50[order(sigtab_unplanted_miPEPc_top_50$VALUE, na.last=NA), ]

my_labels<-c(sigtab_unplanted_miPEPc_order$Genus)

plot_miPEPc_unplanted_order <- plot_miPEPc_unplanted + scale_x_discrete(
  labels=my_labels) + scale_color_manual(values = c("darkolivegreen1","darkorchid", "darkblue","brown1","deeppink","firebrick","darksalmon","darkcyan","darkorange","turquoise","magenta","green"))
plot_miPEPc_unplanted_order <- plot_miPEPc_unplanted_order + labs(x = "Genus") + coord_flip() + labs(y="shrunken log2FoldChanges")
plot_miPEPc_unplanted_order

rm(phylo_unplanted_C)
```

4) Etienne's -- make list of ASVs that are affected in root, but not in unplanted -- through plant effect
```{r}
#miPEP-A
row.names(sigtab_roots_miPEPa)[row.names(sigtab_roots_miPEPa) %in% row.names(sigtab_unplanted_miPEPa)] #642
phylo_roots@tax_table["642",] #Taxonomy: "Bacteria" "Proteobacteria" "Alphaproteobacteria" "Rhizobiales" "KF-JG30-B3" "KF-JG30-B3FA"

#miPEP-B
row.names(sigtab_roots_miPEPb)[row.names(sigtab_roots_miPEPb) %in% row.names(sigtab_unplanted_miPEPb)] #582, 940, 1347, 1197, 1155, 1000
phylo_roots@tax_table["582",] #"Bacteria" "Bacteroidetes" "Sphingobacteriia" "Sphingobacteriales" "Chitinophagaceae" "Taibaiella" 
phylo_roots@tax_table["940",] #"Bacteria" "Bacteroidetes" "Sphingobacteriia" "Sphingobacteriales" "Sphingobacteriaceae" "Pedobacter"
phylo_roots@tax_table["1347",] #"Bacteria" "Planctomycetes" "Phycisphaerae" "Tepidisphaerales" "Tepidisphaeraceae" "TepidisphaeraceaeFA"
phylo_roots@tax_table["1197",] #"Bacteria" "Proteobacteria" "Alphaproteobacteria" "Rhodospirillales" "Rhodospirillaceae" "Dongia" 
phylo_roots@tax_table["1155",] #"Bacteria" "Actinobacteria" "Actinobacteria" "Frankiales" "Acidothermaceae" "Acidothermus"
phylo_roots@tax_table["1000",] #"Bacteria" "Acidobacteria" "Acidobacteria" "Acidobacteriales" "Acidobacteriaceae (Subgroup 1)", "Granulicella"

#miPEP-C
row.names(sigtab_roots_miPEPc)[row.names(sigtab_roots_miPEPc) %in% row.names(sigtab_unplanted_miPEPc)] #530, 1060
phylo_roots@tax_table["530",] #"Bacteria" "Proteobacteria" "Gammaproteobacteria" "Xanthomonadales" "Xanthomonadales Incertae Sedis", "Steroidobacter"
phylo_roots@tax_table["1060",] #"Bacteria" "Bacteroidetes" "Sphingobacteriia" "Sphingobacteriales" "Saprospiraceae" "Lewinella"

#Overlap between miPEPs
row.names(sigtab_roots_miPEPa)[row.names(sigtab_roots_miPEPa) %in% row.names(sigtab_roots_miPEPb)] #"1083" "267"  "642"  "1228" "414"  "790"  "249"  "582"  "576"  "2836" "990"  "620"  "768"  "700"  "1563"
row.names(sigtab_roots_miPEPa)[row.names(sigtab_roots_miPEPa) %in% row.names(sigtab_roots_miPEPc)] #"267"  "642"  "1228" "790"  "422"  "872"  "520"  "1074" "914"  "1994"
row.names(sigtab_roots_miPEPb)[row.names(sigtab_roots_miPEPb) %in% row.names(sigtab_roots_miPEPc)] #"753"  "267"  "474"  "64"   "297"  "614"  "173"  "1228" "333"  "790"  "642"  "453"  "417"  "839"  "309" "1347" "816"  "1000" "1398" "1162" "1639" "894" 


#Venn diagram
library(gplots)
#Root
root.venn <-venn(list(row.names(sigtab_roots_miPEPa), 
     row.names(sigtab_roots_miPEPb),
     row.names(sigtab_roots_miPEPc))
     )
attr(root.venn, "intersections")$`A:B:C` #267, 642, 1228,790

#Shared taxonomy
phylo_roots@tax_table["267",]
phylo_roots@tax_table["642",]
phylo_roots@tax_table["790",]
phylo_roots@tax_table["1228",]

#unplanted
unplanted.venn <- venn(list(row.names(sigtab_unplanted_miPEPa), 
     row.names(sigtab_unplanted_miPEPb),
     row.names(sigtab_unplanted_miPEPc))
     )
attr(unplanted.venn, "intersections")$`A:B:C` #750, 752, 865, 1988, 2735

#Shared taxonomy
phylo_unplanted@tax_table["750",]
phylo_unplanted@tax_table["752",]
phylo_unplanted@tax_table["865",]
phylo_unplanted@tax_table["1988",]
phylo_unplanted@tax_table["2735",]

```

Indicator Species

1) Roots samples
```{r}
#library(indicspecies)

# Export the OTU table, with a Sample column, Type column and then ASV & abundances per column
#OTU_table_phylo_roots <- otu_table(phylo_roots)
#head(OTU_table_phylo_roots)
#write.table(OTU_table_phylo_roots, "OTU_table_phylo_roots.txt")

#Re-import the table OTU_table_phylo_roots_1.xlsx
#OTU_table_phylo_roots_1<-read_excel("C:/Users/hmiddleton/Dropbox/Projet_Mutants_miPEPs_miRNA/Harriet/Metabarcoding_mutants_and_miPEP/Analyses_métabarcoding/miPEP_H/16S/OTU_table_phylo_roots_1.xlsx")

#abund_roots = OTU_table_phylo_roots_1[,3:ncol(OTU_table_phylo_roots_1)] #ASV abundance starts at column 3
#Type = OTU_table_phylo_roots_1$Type # create a vector that contains the information from the "Type" column
#set.seed(771)
#inv_roots = multipatt(abund_roots, Type, func = "r.g", control = how(nperm=9999))
#summary(inv_roots)
```

2) Soil samples
```{r}
#https://jkzorz.github.io/2019/07/02/Indicator-species-analysis.html
#library(indicspecies)

# Export the OTU table, with a Sample column, Type column and then ASV & abundances per column
#OTU_table_phylo_soil <- otu_table(phylo_soil)
#head(OTU_table_phylo_soil)
#write.table(OTU_table_phylo_soil, "OTU_table_phylo_soil.txt")

#Re-import the table OTU_table_phylo_soil_1.xlsx
#OTU_table_phylo_soil_1<-read_excel("C:/Users/hmiddleton/Dropbox/Projet_Mutants_miPEPs_miRNA/Harriet/Metabarcoding_mutants_and_miPEP/Analyses_métabarcoding/miPEP_H/16S/OTU_table_phylo_soil_1.xlsx")

#abund_soil = OTU_table_phylo_soil_1[,3:ncol(OTU_table_phylo_soil_1)] #ASV abundance starts at column 3
#Type = OTU_table_phylo_soil_1$Type # create a vector that contains the information from the "Type" column
#set.seed(771)
#inv_soil = multipatt(abund_soil, Type, func = "r.g", control = how(nperm=9999))
#summary(inv_soil)
```

3) Unplanted samples
```{r}
#library(indicspecies)

# Export the OTU table, with a Sample column, Type column and then ASV & abundances per column
#OTU_table_phylo_unplanted <- otu_table(phylo_unplanted)
#head(OTU_table_phylo_unplanted)
#write.table(OTU_table_phylo_unplanted, "OTU_table_phylo_unplanted.txt")

#Re-import the table OTU_table_phylo_unplanted_1.xlsx
#OTU_table_phylo_unplanted_1<-read_excel("C:/Users/hmiddleton/Dropbox/Projet_Mutants_miPEPs_miRNA/Harriet/Metabarcoding_mutants_and_miPEP/Analyses_métabarcoding/miPEP_H/16S/OTU_table_phylo_unplanted_1.xlsx")

#abund_unplanted = OTU_table_phylo_unplanted_1[,3:ncol(OTU_table_phylo_unplanted_1)] #ASV abundance starts at column 3
#Type = OTU_table_phylo_unplanted_1$Type # create a vector that contains the information from the "Type" column
#set.seed(771)
#inv_unplanted = multipatt(abund_unplanted, Type, func = "r.g", control = how(nperm=9999))
#summary(inv_unplanted)
```


Etienne's simplified graphs -- community composition
```{r}
pseq_roots <- phylo_roots %>% aggregate_taxa(level = "Phylum")
pseq_unplanted <- phylo_unplanted %>% aggregate_taxa(level = "Phylum")

keep_phyla <- names(sort(taxa_sums(pseq_roots), decreasing = TRUE))[1:10]

##Roots
# Transform Taxa counts to relative abundance using total number of reads - include others
ps1_roots_phylum_relabun_EY <- data.frame(apply(pseq_roots@otu_table, 1, "/", colSums(pseq_roots@otu_table))*100)
rowSums(ps1_roots_phylum_relabun_EY) #100 everywhere
ps1_roots_phylum_relabun_EY_top10 <- ps1_roots_phylum_relabun_EY[,colnames(ps1_roots_phylum_relabun_EY)%in%keep_phyla]
others <- 100-rowSums(ps1_roots_phylum_relabun_EY_top10)
ps1_roots_phylum_relabun_EY_top10 <- data.frame(ps1_roots_phylum_relabun_EY_top10, others)
rowSums(ps1_roots_phylum_relabun_EY_top10)#Sanity check; 100 all over

#Add mapping file info
sample.s <- sample[order(row.names(sample)),]#Sort mapping file
sample.root.s <- sample.s[sample.s$Compartment=="RACINE",]
sample.root.s <- sample.root.s[-23,] #One sample that was removed from the OTU table
ps1_roots_phylum_relabun_EY_top10.s <- ps1_roots_phylum_relabun_EY_top10[order(row.names(ps1_roots_phylum_relabun_EY_top10)),] #Sort Phylum table
row.names(ps1_roots_phylum_relabun_EY_top10.s)==row.names(sample.root.s)#Sanity check - all TRUE
phy.map.root <- data.frame(sample.root.s,ps1_roots_phylum_relabun_EY_top10.s)#Merge
rowSums(phy.map.root[,5:15]) #Should be all 100s
phy.map.root.long <- gather(phy.map.root,Phylum,relabund,5:15) #transform in long format for ggplot
#tax.map.long$growthstage <- factor(tax.map.long$growthstage, c("tillering", "stemelongation", "booting", "heading", "flowering" ))#Reorder manually the growth stages

##Unplanted
# Transform Taxa counts to relative abundance using total number of reads - include others
ps1_unplanted_phylum_relabun_EY <- data.frame(apply(pseq_unplanted@otu_table, 1, "/", colSums(pseq_unplanted@otu_table))*100)
rowSums(ps1_unplanted_phylum_relabun_EY) #100 everywhere
ps1_unplanted_phylum_relabun_EY_top10 <- ps1_unplanted_phylum_relabun_EY[,colnames(ps1_unplanted_phylum_relabun_EY)%in%keep_phyla]
others <- 100-rowSums(ps1_unplanted_phylum_relabun_EY_top10)
ps1_unplanted_phylum_relabun_EY_top10 <- data.frame(ps1_unplanted_phylum_relabun_EY_top10, others)
rowSums(ps1_unplanted_phylum_relabun_EY_top10)#Sanity check; 100 all over

#Add mapping file info
sample.unplanted.s <- sample.s[sample.s$Compartment=="SOL.NU",]
ps1_unplanted_phylum_relabun_EY_top10.s <- ps1_unplanted_phylum_relabun_EY_top10[order(row.names(ps1_unplanted_phylum_relabun_EY_top10)),] #Sort Phylum table
row.names(ps1_unplanted_phylum_relabun_EY_top10.s)==row.names(sample.unplanted.s)#Sanity check - all TRUE
phy.map.unplanted <- data.frame(sample.unplanted.s,ps1_unplanted_phylum_relabun_EY_top10.s)#Merge
rowSums(phy.map.unplanted[,5:15]) #Should be all 100s
phy.map.unplanted.long <- gather(phy.map.unplanted,Phylum,relabund,5:15) #transform in long format for ggplot

#Merge both long tables
phylum.long <- rbind(phy.map.root.long,phy.map.unplanted.long)

#To have correct label names
comp.labs <- c("Roots","Unplanted")
names(comp.labs) <- c("RACINE", "SOL.NU")

StackedBarPlot_phylum_rel_EY <- ggplot(phylum.long, aes(x =Type, y = relabund, fill = Phylum)) +
  geom_bar(stat = "summary", fun ="mean", position = "stack") +
  labs(y = "Relative abundance (%)", x="Treatment") +
  facet_grid(~ Compartment, scales = "free", labeller = labeller(Compartment = comp.labs)) +
  scale_y_continuous(limits = c(0,100), expand = c(0,0)) +
  scale_x_discrete(labels=c("Water", "miPEP-A", "miPEP-B", "miPEP-C"))+
  theme_bw()+
scale_fill_manual(values = c("darkblue", "darkgoldenrod1","darkgrey", "darkseagreen", "darkorchid", "darkolivegreen1", "lightskyblue", "darkgreen", "deeppink", "khaki2", "firebrick"))

StackedBarPlot_phylum_rel_EY


```


Etienne's simplified graphs -- alpha diversity
```{r}
div.table <- rbind(p_Div_roots$data, p_Div_unplanted$data)
div.table.shannon <- div.table[div.table$variable=="Shannon",]

# Test normality
shapiro.test(div.table.shannon[div.table.shannon$Compartment=="RACINE",]$value) #W = 0.9674, p-value = 0.4506
shapiro.test(div.table.shannon[div.table.shannon$Compartment=="SOL.NU",]$value) #W = 0.89464, p-value = 0.005312

#Anova
anova.div.root <- aov(value~Type, data = div.table.shannon[div.table.shannon$Compartment=="RACINE",])
anova.div.unplanted <- aov(value~Type, data = div.table.shannon[div.table.shannon$Compartment=="SOL.NU",])

summary(anova.div.root) #F=1.292, P=0.297
summary(anova.div.unplanted) #F=6.587. P=0.00174
TukeyHSD(anova.div.unplanted) #ctrl: a, A: b, B: a, C:a
tuk.labels.div <- data.frame(x=c(1,2,3,4), y=c(rep(5.59, 4)), Compartment="SOL.NU", symbol=c("a","b","a","a")) #To add letters on the graph

#To have correct label names
comp.labs <- c("Roots","Unplanted")
names(comp.labs) <- c("RACINE", "SOL.NU")

div.table.shannon$Type <- factor(div.table.shannon$Type, c("Control-mi-PEP", "miPEP-A", "miPEP-B", "miPEP-C"))#Reorder manually the treatments

p_Div <- ggplot(div.table.shannon, aes(x=Type, y=value))+
    geom_boxplot()+ 
    geom_point()+
    geom_text(data = tuk.labels.div, aes(x=x, y=y, label=symbol), size=5)+
    theme_bw()+
    facet_grid(~Compartment, labeller = labeller(Compartment = comp.labs))+
    xlab("Treatment")+
    ylab("Shannon diversity index")+ 
    scale_x_discrete(labels=c("Water", "miPEP-A", "miPEP-B", "miPEP-C"))+
  #theme(axis.text.y = element_text (), axis.text.x = element_blank (), axis.text =element_text(size=8), axis.title=element_text(size=12,face="bold"))+
    scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02", "#A6761D","#666666"))

p_Div


```


qPCR graphs
```{r}
qPCR <- read.table(here("data", "qpcr.txt"), header = T, sep="\t")
qPCR.s <- qPCR[order(qPCR$Sample),]
qPCR.s$Sample==sample.s$X.SampleID #Sort check: 96 TRUE
qPCR.sample <- cbind(sample.s,qPCR.s)

# Test normality
shapiro.test(qPCR.sample[qPCR.sample$Compartment=="RACINE",]$X16S) #W = 0.9264, p-value = 0.03112
shapiro.test(qPCR.sample[qPCR.sample$Compartment=="SOL.NU",]$X16S) #W = 0.97239, p-value = 0.5679

#Anova
anova.16S.root <- aov(X16S~Type, data = qPCR.sample[qPCR.sample$Compartment=="RACINE",])
anova.16S.unplanted <- aov(X16S~Type, data = qPCR.sample[qPCR.sample$Compartment=="SOL.NU",])

summary(anova.16S.root) #F=1.94, P=0.146
summary(anova.16S.unplanted) #F=4.84. P=0.00775
TukeyHSD(anova.16S.unplanted) #ctrl: a, A: b, B: ab, C:ab
tuk.labels.16S <- data.frame(x=c(1,2,3,4), y=c(rep(2e+05, 4)), Compartment="SOL.NU", symbol=c("a","b","ab","ab")) #To add letters on the graph

#To have correct label names
comp.labs <- c("Roots","Unplanted")
names(comp.labs) <- c("RACINE", "SOL.NU")

#16S
bact.qPCR.box <- ggplot(qPCR.sample[qPCR.sample$Compartment!="SOL",], aes(x=Type, y=X16S)) +
  geom_boxplot() +
  geom_point()+
  geom_text(data = tuk.labels.16S, aes(x=x, y=y, label=symbol), size=5)+
  facet_grid(~Compartment, labeller = labeller(Compartment = comp.labs))+
  theme_bw() + 
  ylab("16S rRNA gene copies (per ng DNA)") + theme(legend.position="none")+
  xlab("Treatment")+
  scale_x_discrete(labels=c("Water", "miPEP-A", "miPEP-B", "miPEP-C"))+
  scale_y_log10()
bact.qPCR.box

```

Put graphs together in one figure
```{r}
library(ggpubr)

Fig1 <- ggarrange(StackedBarPlot_phylum_rel_EY,p_Div,bact.qPCR.box, ncol = 1, legend = "bottom", labels = c("A","B","C"))
Fig1
ggsave(filename = here("output", "Fig1.tiff"),  plot = Fig1, compression = "lzw", height = 14, units = "in")

```